Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
add better algorithms descriptions and algs

First, \PCF{CannotMap} checks if there could possibly be a lens satisfying the
examples. If there is not, \GreedySynth returns \None early. After this,
\GreedySynth finds the best lenses between each sequence pair of the left and
right DNF regular expressions, given the examples that match them. A priority
queue containing these sequence lenses is initialized from these
($\PCF{PQ.Create}$). The symmetric lens is built up iteratively from these
sequence lenses, the state of this partial lens is tracked in the lens builder,
$lb$.

\GreedySynth then loops until there are no more sequence lenses in the priority
queue. Within this loop, sequence lenses are popped. If a popped sequence lens
maps between two sequences which aren't yet included in a sequence lens, then
this sequence lens is added to the partial DNF lens. Furthermore, we update the
priorities of our sequence lenses as the algorithm proceeds: if two sequence
lenses have the same source, the second one to be popped gets a higher cost than
it originally had, to address the information that needs to be stored for
including that source of non-bijectivity.

In our example, \GreedySynth may first pop the sequence pair \lstinline{""} and
\lstinline{""}, as there is a bijective sequence lens between them. After this,
the sequence lens between \lstinline{name.name*} and \lstinline{name} would be
popped. While examples prevent lenses between the other sequence pairs, if there
were other lenses, they would be popped, but not added to the ongoing state, as
they would not be a useful addition to the state.

Finally, after all sequences have been
popped, the partial DNF lens is converted into a symmetric DNF lens. This is
only possible if all sequences are involved in some sequence lens: if they are
not, \PCF{GreedyState.ToDNFLens} instead returns \None.

\begin{algorithm}
  \caption{\GreedySynth}
  \label{alg:greedysynth}
  \begin{algorithmic}[1]
    \Function{GreedySynth}{$\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}$}
    \If{$\Call{CannotMap}{\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\SLS \gets
    \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
    \State $\GreedyState \gets \Call{PQ.Create}{\SLS}$
    \State $lb \gets \Call{LensBuilder.Empty}{}$
    \While{$\Call{PQ.IsNonempty}{\GreedyState}$}
    \State $\SSQLens \gets
    \Call{PQ.PopBest}{\GreedyState}$
    \If{$\Call{LensBuilder.UsefulAdd}{lb,\SSQLens}$}
    \State $lb \gets \Call{GreedyState.AddSeq}{lb,\SSQLens}$
    \EndIf
    \EndWhile
    \State \ReturnVal{\Call{LensBuilder.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

This algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom
lenses by iterating the DNF lens between its subcomponents. The inputs to
\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
between the given sequences regular expressions that is consistent with the
examples. If there is not, \GreedySeqSynth returns \None early. After this,
\GreedySeqSynth finds the best lenses between each atom pair of the left and
right sequences. An internal state for the algorithm is initialized from these.
This state consists of two things: an intermediate sequence lens and a priority
queue consisting of potential sequence lenses. On initialization,
\GreedySeqSynth throws all the atom lenses into the priority queue, and the
intermediate sequence lens is one that projects all the atoms.

For example, if trying to find a sequence lens between \lstinline{name.name*}
and \lstinline{name}, lenses are found between \lstinline{name} and
\lstinline{name}, and \lstinline{name} and \lstinline{name*}, and added to the
queue.

\GreedySeqSynth then loops until there are no more atom lenses in the priority
queue. Within this loop, atom lenses are popped. The atom lens is then added if
it is either necessary, or useful. Within a sequence lens, each atom can either
be projected or be synchronized with exactly one atom in the other sequence, and
examples can show that certain atoms \emph{need} to be mapped. In such
situations, that atom lens is necessary. A useful atom lens is where adding that
lens reduces the overall cost of the partial lens. If both atoms are not
involved in an atom lens, the lens will always be useful, but sometimes adding
that lens requires removing another lens. If the cost added by removing that
other lens is more than is removed by adding the proposed lens, then this atom
lens is not included in the sequence lens. For example, the bijective lens
between \lstinline{name} and \lstinline{name} would first be found, and added to
the state.  Next, the disconnect lens between \lstinline{name} and
\lstinline{name*} would be popped.  Because this lens is not useful, it is not
added to the state.

Finally, after all atoms have been
popped, the sequence lens is validated and returned. The sequence lens is
validated if all atoms that need to be mapped, are: if they are not,
\PCF{GreedySeqState.ToDNFLens} instead returns \None.

@@

    \Function{GreedySeqSynth}{$\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}$}
    \If{$\Call{CannotMap}{\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\ALS \gets
    \Call{\CartesianMap}{\AtomSynth(\Examples),\ListOf{\Atom_1;\ldots;\Atom_n},\ListOf{\AtomAlt_1;\ldots;\AtomAlt_m}}$
    \State $\GreedyState \gets \Call{GreedySeqState.Init}{\ALS}$
    \While{$\Call{GreedySeqState.Continue}{\GreedyState}$}
    \State $(\GreedyState,\SAtomLens) \gets
    \Call{GreedySeqState.PopBest}{\GreedyState}$
    \If{$\Call{GreedySeqState.UsefulAdd}{\GreedyState,\SAtomLens}
      \BooleanOr \Call{NecessaryAdd}{\Examples,\SAtomLens}$}
    \State $\GreedyState \gets \Call{GreedySeqState.AddAtom}{\GreedyState,\SAtomLens}$
    \EndIf
    \EndWhile
    \State \ReturnVal{\Call{GreedySeqState.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

>>>>

First, \PCF{CannotMap} checks if there could possibly be a lens satisfying the
examples. If there is not, \GreedySynth returns \None early. After this,
\GreedySynth finds the best lenses between each sequence pair of the left and
right DNF regular expressions, given the examples that match them. A priority
queue containing these sequence lenses is then initialized with
$\PCF{PQ.Create}$. The symmetric lens is built up iteratively from these
sequence lenses, the state of this partial lens is tracked in the lens builder,
$lb$.

\GreedySynth then loops until there are no more sequence lenses in the priority
queue. Within this loop, sequence lenses are popped. If this sequence lens is
``useful'', then that sequence lens is included in the final DNF lens. If the
source or target of a popped sequence lens is not the source or target of
another sequence lens, then this sequence lens is considered useful. If the
examples require two sequences have a lens between them, then any sequence lens
between them is considered useful. We update the priorities of our sequence
lenses as the algorithm proceeds: if two sequence lenses have the same source,
the second one to be popped gets a higher cost than it originally had, to
address the information that needs to be stored for including that source of
non-bijectivity.

In our example, \GreedySynth may first pop the sequence pair \lstinline{""} and
\lstinline{""}, as there is a bijective sequence lens between them. As neither
\lstinline{""} is involved in a sequence lens, this lens is considered useful.
After this, the sequence lens between \lstinline{name.name*} and
\lstinline{name} would be popped: while that lens is not bijective: it is still
better than the alternatives. As all sequences are now involved in sequence lenses,
and there are no examples to make other lenses useful, no further sequence
lenses would be added to the lens builder.

Finally, after all sequences have been popped, the partial DNF lens $lb$ is
converted into a symmetric DNF lens. This is only possible if all sequences are
involved in some sequence lens: if they are not, \PCF{GreedyState.ToDNFLens}
instead returns \None.

\begin{algorithm}
  \caption{\GreedySynth}
  \label{alg:greedysynth}
  \begin{algorithmic}[1]
    \Function{GreedySynth}{$\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}$}
    \If{$\Call{CannotMap}{\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\SLS \gets
    \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
    \State $\GreedyState \gets \Call{PQ.Create}{\SLS}$
    \State $lb \gets \Call{LensBuilder.Empty}{}$
    \While{$\Call{PQ.IsNonempty}{\GreedyState}$}
    \State $\SSQLens \gets
    \Call{PQ.Pop}{\GreedyState}$
    \If{$\Call{LensBuilder.UsefulAdd}{lb,\SSQLens,\Examples}$}
    \State $lb \gets \Call{GreedyState.AddSeq}{lb,\SSQLens}$
    \EndIf
    \EndWhile
    \State \ReturnVal{\Call{LensBuilder.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

This algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom
lenses by iterating the DNF lens between its subcomponents. The inputs to
\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
between the given sequences regular expressions that is consistent with the
examples. If there is not, \GreedySeqSynth returns \None early. After this,
\GreedySeqSynth finds the best lenses between each atom pair of the left and
right sequences. A priority queue containing these lenses is then initialized
with \PCF{PQ.Create}. The symmetric sequence lens is built up iteratively from
these atom lenses, the state of the partial lens is tracked in the sequence lens
builder, $slb$.

\GreedySeqSynth then loops until there are no more atom lenses in the priority
queue. Within this loop, atom lenses are popped. The atom lens is then added if
it is useful. An atom lens is useful if adding it will lower the cost of the
generated sequence lens, or if examples show that one of the atoms must be
connected to another lens. The priorities of these lenses change over time, as
each atom can only be involved in one lens. Adding a lens may necessitate
removing another, so not all popped atom lenses will lower the cost of the
generated sequence lens.

Finally, after all atoms have been popped, the sequence lens is validated and
returned. The algorithm confirms that all atoms that need to be connected with
an atom lens, are. If they are not, \PCF{SLensBuilder.ToDNFLens} instead returns
\None.

@@

    \Function{GreedySeqSynth}{$\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}$}
    \If{$\Call{CannotMap}{\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\ALS \gets
    \Call{\CartesianMap}{\AtomSynth(\Examples),\ListOf{\Atom_1;\ldots;\Atom_n},\ListOf{\AtomAlt_1;\ldots;\AtomAlt_m}}$
    \State $\GreedyState \gets \Call{PQ.Init}{\ALS}$
    \State $slb \gets \Call{SLensBuilder.Empty}{}$
    \While{$\Call{PQ.IsNonempty}{\GreedyState}$}
    \State $\SAtomLens \gets
    \Call{PQ.Pop}{\GreedyState}$
    \If{$\Call{SLensBuilder.UsefulAdd}{slb,\SAtomLens,\Examples}$}
    \State $\GreedyState \gets \Call{SLensBuilder.AddAtom}{\GreedyState,\SAtomLens}$
    \EndIf
    \EndWhile
    \State \ReturnVal{\Call{SLensBuilder.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}
